SDSS-IV MaNGA: Mass Independent Star Formation Enhancement in Close Galaxy Pairs

### 2020-07-28 maintex.tex
edits based on Hai's suggestions in Sections 1-2

\title{SDSS-IV MaNGA: The Universal Profile of Enhanced Star Formation in Close Galaxy Pairs}

@@

\begin{abstract}
We study the spatial profile of the star formation rate (SFR) in a sample of 160 interacting galaxies within the MaNGA (Mapping Nearby Galaxies at Apache Point Observatory) survey. We find that the specific star formation rate (sSFR) is centrally (e.g. within a single effective radius) enhanced in interacting galaxies by $\sim$0.25 $\pm$ 0.1 but that there is no enhancement or suppression in their disks. We split the radial profile into four separate stellar mass bins from \logm\ $=$ 9.0$-$11.0 and find that the sSFR enhancement has no dependence on the stellar mass of the paired galaxies. Within one effective radius, the sSFR enhancement in each of the mass bins falls within $\sim$0.1 dex of the median profile for the whole pair sample. Low mass galaxies experience the same level of enhancement to their star formation rate as high mass galaxies. Assuming a constant burst merger induced star formation over a 1 Gyr period, the mass growth rate in the centers of the paired galaxies will be 1.6$-$2.3$\times$ higher than secular mass growth. Since the merger induced sSFR enhancement is constant over stellar mass, low mass galaxies will experience a higher fractional mass growth rate which is $\sim$4.3 higher than that of high mass galaxies. This indicates that merger induced star formation has a significant impact on the mass evolution of low mass galaxies.

@@

%%%Observations%%%
Centrally enhanced star formation has also been observed in galaxy pairs with large spectroscopic surveys including; CfA2  \citep{Barton:2000, Woods:2006}, 2dF \citep{Lambas:2003}, AEGIS \citep{Lin:2007}, SDSS \citep{Ellison:2008}, COSMOS \citep{Kartaltepe:2007,Xu:2012}, PRIMUS \citep{Wong:2011}, and GAMA \citep{Robotham:2014}.

@@

% Organization
This paper is organized as follows; in Section \ref{sec:data} we will discuss the properties of the MaNGA survey along with the construction of our pair and control samples, in Section \ref{sec:analysis} we will discuss how we measure star formation rates and how we build radial profiles of star formation, in Section \ref{sec:results} we discuss the results of the constructed radial profiles, and in Section \ref{sec:sum} we discuss and summarize the findings of the work.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
\centering
\includegraphics[width=\linewidth]{fig/color-mag.pdf}
\caption[]{Color-magnitude diagram for MaNGA galaxies ({\it colored circles}). The color of the symbol reflects the local density around each data point in this color-magnitude plane, as indicated by the color bar on the top. The MaNGA galaxies with close companions are marked with {\it grey diamonds}. From top to bottom, the dashed lines divide the sample into red sequence, green valley, and blue cloud. The star-forming galaxy sample used in this paper are the galaxies below the lower dividing line.}
\label{fig:cmd}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

The MaNGA survey targets galaxies from a subset of 41,154 galaxies from the NASA-Sloan Atlas (NSA) with a redshift range of 0.01 $<$ z $<$ 0.15 and a luminosity range of -17.7 $<$ $\mathcal{M_{\rm i}}$ $<$ -24.0, where $\mathcal{M_{\rm i}}$ is the rest frame i-band magnitude within the survey's elliptical Petrosian apertures. MaNGA plans to cover 10,000 galaxies with a flat stellar mass distribution at two spatial coverages, 1.5 R/\reff\ and 2.5 R/\reff\ (where \reff\ is the radius which contains 50\% of the galaxy's total light). In this work we use the data from the internal data release, MPL-8, which covers 6142 unique galaxies.

@@

\begin{equation}\label{eq:blue}
NUV-r = 3.1682 - 0.16 (M_i+18)
\end{equation}
\begin{equation}\label{eq:red}
NUV-r = 4.7866 - 0.04 (M_i+18)
\end{equation}

@@

After the star formation cuts, we require that all galaxies are in the Primary or Secondary MaNGA subsamples \citep{Wake:2017}. We also set a stellar mass range of \logm\ $=$ 9.0$-$11.0.

@@

In our previous work, \citet{Fu:2018}, we showed that $\sim$5.7\% of the MaNGA observations have at least one companion galaxy within the field of view of the IFU. In this work we use a refined pair identification process which is also expanded for the most recent internal data release, MPL-8.

@@

We use the r-band elliptical Petrosian apertures from the NSA catalog to describe the geometry of the galaxies in the outside-IFU sample. For cases where the identified galaxy pair is outside of the MaNGA IFU, the NSA catalog's algorithm is able to fit the MaNGA target's geometry well. We also use the stellar masses calculated from the r-band elliptical Petrosian apertures in the NSA catalog for the total stellar mass of the galaxies.

@@

We use the BPT diagnostic \citep{Baldwin:1981} to remove spaxels which were not ionized by star formation. Ionization sources are often separated by either the theoretical maximum starburst line from \citet{Kewley:2001} or the analytically fitted line from \citet{Kauffmann:2003}, given below in Equations \ref{eq:kewley} and \ref{eq:kauffmann} respectively.

\begin{equation}\label{eq:kewley}
\rm{log\Big(\frac{[O\,\textsc{iii}]\lambda 5007}{H\beta}\Big) = \frac{0.61}{log([N\,\textsc{ii}]\lambda 6584 / H\alpha) - 0.47} + 1.19}
\end{equation}
\begin{equation}\label{eq:kauffmann}
\rm{log\Big(\frac{[O\,\textsc{iii}]\lambda 5007}{H\beta}\Big) = \frac{0.61}{log([N\,\textsc{ii}]\lambda 6584 / H\alpha) - 0.05} + 1.3}
\end{equation}

Where \OIII$\lambda$5007, H$\beta$, \NII$\lambda$6584, and H$\alpha$ are emission line fluxes within each MaNGA spaxel. We depict the BPT diagram for each spaxel in the MaNGA survey in Figure \ref{fig:bpt} along with the demarcations for star formation from \citet{Kewley:2001} and \citet{Kauffmann:2003}. We use the \citet{Kewley:2001} demarcation line as the limit for classifying star formation in the MaNGA spaxels.

@@

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Discussion and Summary}\label{sec:sum}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
\centering
\includegraphics[width=3in]{fig/mass_gain.pdf}
\caption[The fractional mass gain due to merger induced star formation.]{The fractional mass growth due to merger induced star formation over a dynamical timescale, $\tau$ $=$ 1 Gyr. The highlighted region represents the propagated standard errors of the mean of the stacked log(sSFR) profiles. }
\label{fig:mass_gain_sum}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

Looking at individual difference profiles in our work, we see that there is a large variance in the $\Delta$log(sSFR) where there are as many paired galaxies with sSFR enhancement in their disks as there are paired galaxies with suppression in there disks. The end result is that the average of the $\Delta$log(sSFR) is zero. The merger induced effects seen in the disks of the paired galaxies may be more complicated than what is seen in the centers of the galaxies. \\

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\section{Summary and Conclusion}\label{sec:sum}

>>>>

\title{
SDSS-IV MaNGA: Mass Independent Star Formation Enhancement in Close Galaxy Pairs
}

@@

\begin{abstract}
We study the spatial profile of the star formation rate (SFR) in a sample of {\it 160} interacting galaxies within the SDSS-IV MaNGA survey. We find that the specific star formation rate (sSFR) is centrally (e.g. within a single effective radius) enhanced in interacting galaxies by $\sim$0.25 $\pm$ 0.1 but that there is no enhancement or suppression in their disks. We split the radial profile into five separate stellar mass bins from \logm\ $=$ 9.0$-$11.5 and find that the sSFR enhancement has no dependence on the stellar mass of the paired galaxies. Within one effective radius, the sSFR enhancement in each of the mass bins falls within $\sim$0.1 dex of the median profile for the whole pair sample. We then use a second method to compare paired galaxies and control galaxies where each paired galaxies is matched to a set of 20 control galaxies based on stellar mass and redshift. The sSFR enhancements found are still mass independent, reinforcing the idea that low mass galaxies experience the same level of enhancement to their star formation rate as high mass galaxies. Since the merger induced sSFR enhancement is constant over stellar mass, low mass galaxies will experience a higher fractional mass growth rate of 1$-$7\%, assuming a constant burst merger induced star formation over a 1 Gyr period. This indicates that merger induced star formation has a significant impact on the mass evolution of low mass galaxies.

@@

%%%Observations%%%
Centrally enhanced star formation has also been observed in galaxy pairs with large spectroscopic surveys including; CfA2 (Center of Astrophysics Redshift Survey)  \citep{Barton:2000, Woods:2006}, 2dF (Two Degree Field) \citep{Lambas:2003}, AEGIS (All-Wavelength Extended Groth Strip International Survey) \citep{Lin:2007}, SDSS (Sloan Digital Sky Survey) \citep{Ellison:2008}, COSMOS (Cosmic Evolution Survey) \citep{Kartaltepe:2007,Xu:2012}, PRIMUS (Prism Multi-Object Survey) \citep{Wong:2011}, and GAMA (Galaxy and Mass Assembly) \citep{Robotham:2014}.

@@

% Organization
This paper is organized as follows; in Section \ref{sec:data} we will discuss the properties of the MaNGA survey along with the construction of our pair and control samples, in Section \ref{sec:analysis} we will discuss how we measure star formation rates and how we build radial profiles of star formation, in Section \ref{sec:results} we discuss the mass independence of the constructed radial profiles, in Section \ref{sec:disc} we explore the radial profiles as a function of the mass ratio and projected separation and we compare our work against previous works, and in Section \ref{sec:sum} we summarize the findings of the work.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
\centering
\includegraphics[width=\linewidth]{fig/color-mag.pdf}
\caption[]{Color-magnitude diagram for MaNGA galaxies ({\it colored circles}). The color of the symbol reflects the local density around each data point in this color-magnitude plane, as indicated by the color bar on the top. Inside and outside-IFU pairs are marked with {\it grey diamonds}. From top to bottom, the dashed lines divide the sample into red sequence, green valley, and blue cloud. The star-forming galaxy sample used in this paper are the galaxies below the lower dividing line.}
\label{fig:cmd}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

The MaNGA survey targets galaxies from a subset of 41,154 galaxies from the NASA-Sloan Atlas (NSA v1\_0\_1; \url{http://www.nsatlas.org}) with a redshift range of 0.01 $<$ z $<$ 0.15 and a luminosity range of -17.7 $<$ $\mathcal{M_{\rm i}}$ $<$ -24.0, where $\mathcal{M_{\rm i}}$ is the rest frame i-band magnitude within the survey's elliptical Petrosian apertures. MaNGA plans to cover 10,000 galaxies with a flat stellar mass distribution at two spatial coverages, 1.5 R/\reff\ and 2.5 R/\reff\ (where \reff\ is the radius which contains 50\% of the galaxy's total light). In this work we use the data from the internal data release, MPL-8, which covers 6142 unique galaxies.

@@

\begin{equation}\label{eq:blue}
NUV-r = 3.1682 - 0.16 (\mathcal{M}_i+18)
\end{equation}
\begin{equation}\label{eq:red}
NUV-r = 4.7866 - 0.04 (\mathcal{M}_i+18)
\end{equation}

@@

After the star formation cuts, we require that all galaxies are in the Primary or Secondary MaNGA subsamples \citep{Wake:2017}. The resulting sample cover a stellar mass range of \logm\ $=$ 9.0$-$11.5.

@@

In our previous work, \citet{Fu:2018}, using the 2168 unique galaxies form DR14 \citep{Abolfathi:2018} we showed that $\sim$5.7\% of the MaNGA observations have at least one companion galaxy within the field of view of the IFU. In this work we use a refined pair identification process which is also expanded for the most recent internal data release, MPL-8.

@@

@@

We use the BPT diagnostic \citep{Baldwin:1981} to remove spaxels which were not ionized by star formation. To separate ionization sources, we use the theoretical maximum starburst line from \citet{Kewley:2001} (Equation \ref{eq:kewley}).

\begin{equation}\label{eq:kewley}
\rm{log\Big(\frac{[O\,\textsc{iii}]\lambda 5007}{H\beta}\Big) = \frac{0.61}{log([N\,\textsc{ii}]\lambda 6584 / H\alpha) - 0.47} + 1.19}
\end{equation}

Where \OIII$\lambda$5007, H$\beta$, \NII$\lambda$6584, and H$\alpha$ are emission line fluxes within each MaNGA spaxel. We use the \citet{Kewley:2001} demarcation line as the limit for classifying star formation in the MaNGA spaxels.

@@

We use the r-band elliptical Petrosian apertures from the NSA catalog to describe the geometry of the galaxies in the outside-IFU sample. For cases where the identified galaxy pair is outside of the MaNGA IFU, the NSA catalog's algorithm is able to fit the MaNGA target's geometry well. We also use the stellar masses calculated from the r-band elliptical Petrosian apertures in the NSA catalog for the total stellar mass of the galaxies. \\

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Discussion}\label{sec:disc}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
\centering
\includegraphics[width=3in]{fig/mass_gain.pdf}
\caption[The fractional mass gain due to merger induced star formation.]{The fractional mass growth due to merger induced star formation over a dynamical timescale, $\tau$ $=$ 1 Gyr. The highlighted region represents the propagated standard errors of the mean of the stacked log(sSFR) profiles. }
\label{fig:mass_gain_sum}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

Looking at individual difference profiles in our work, we see that there is a large variance in the $\Delta$log(sSFR) where there are as many paired galaxies with sSFR enhancement in their disks as there are paired galaxies with suppression in there disks. The end result is that the average of the $\Delta$log(sSFR) is zero. The merger induced effects seen in the disks of the paired galaxies may be more complicated than what is seen in the centers of the galaxies.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Summary and Conclusion}\label{sec:sum}
